---
title: "data_analysis.rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tigris)
library(tidycensus)
library(ggrepel)
library(sf)
```


**Explaining our Data Choices**
We used Stanford Univeristy's data on police traffic stops, specifically in the city of San Diego. We wanted to look at the outcomes of traffic stops, specifically which stops lead to searches, and what searches yield contraband. We are also interested in when and where citations are issued. San Diego provided data on the parameters we were most interested in exploring, as well as a shape file for us to manipulate into a map visualization. From our general knowledge of the area, we knew that San Diego has high Hispanic population and low Black population, so we thought it would be interesting to explore the rates outcomes of different traffic stops by race.

```{r}
san_diego_stops <- read_rds("data/yg821jf8611_ca_san_diego_2020_04_01.rds")
```

This data is already very clean, there are raw rows present indicating how the Stanford team probably changed things. We did not immediately see anything that needed to be cleaned, but when we start working with specific elements of the data, we will have to drop N/A values and assigning value to N/A rows to make our calculations. We should be able to answer all of the questions, we intentionally chose San Diego because it had the data available that we wanted to look at. There are plenty of values, 383,027 stops recorded with 21 different parameters.  

**Framing our Questions, What do we want to know?**

When we decided to look at traffic stops, we were interested in understanding who police officers are most likely pulling over, and what the outcomes of those traffic stops are, but we needed to narrow that focus significantly, because there was a surprising amount of information available to us. We decided to specifically explore age demographics, because age heavily influences insurance rates, and we guessed that younger drivers might also be trusted less by the police. 

*Question 1:*
What age groups are stopped at the highest rate when age is known? 

To form our groups, we chose to loosely base the categories on the ages used by insurance companies to set rates

**Answer**
25 to 35 year old drivers account for the largest portion of people stopped when age is known. The total increases and decreases on a steady curve before and after this category.

```{r}
age_grouped_san_diego <-san_diego_stops %>%
mutate(
  age_group = case_when(
    subject_age < 18 ~ 'under_18',
    subject_age >=18 & subject_age < 25 ~ '18_25',
    subject_age >= 25 & subject_age < 35 ~ '25_35',
    subject_age >= 35 & subject_age < 45 ~ '35_45',
    subject_age >= 45 & subject_age < 55 ~ '45_55',
   subject_age >= 55 & subject_age < 65 ~ '55_65',
   subject_age >= 65 ~'65+'
))

age_grouped_san_diego <- age_grouped_san_diego %>%
  drop_na(age_group)

age_group_counts <- age_grouped_san_diego %>%
  group_by(age_group) %>%
  summarise(count=n())
```


**Question 2:**
Which age group is most often searched for contraband when they are stopped?
**Answer**
5.7% of 18 to 25 year old drivers were searched when stopped, which is 0.6% more than the 25 to 35 age category.

```{r}
searches_age_groups_san_diego <- age_grouped_san_diego %>%
  filter(search_conducted == "TRUE") %>%
  drop_na(age_group) %>%
  group_by(age_group) %>%
  summarise(count=n())

percentage_search <- age_group_counts %>%
    inner_join(searches_age_groups_san_diego, by=c("age_group")) %>%
    rename(total=count.x,search=count.y)

percentage_search <- percentage_search %>%
  mutate(percent_search=(search/total)*100)
```


**Question 3**
When a search is performed, what group most often had contraband? AKA How often were police correct in conducting searches?

**Answer**
11.3% of 45 to 55 year old drivers had contraband found when they were searched, followed by 10.3% of 18 to 25 year old drivers and 10.1% of drivers under 18. The percentage of drivers stopped with contraband found in 25 to 35 is comparatively low with 8.9%

```{r}
searches_age_groups_contraband <- age_grouped_san_diego %>%
    filter(search_conducted == "TRUE", contraband_found == "TRUE") %>%
    group_by(age_group) %>%
    summarise(count=n())

percentage_search_contraband <- percentage_search %>%
    inner_join(searches_age_groups_contraband, by=c("age_group")) %>%
    rename(contraband=count)

percentage_search_contraband <- percentage_search_contraband %>%
  mutate(percent_contraband=(contraband/search)*100)

```



We also knew that later on in our analysis, we would want to consider racial demographics because news coverage has consistently shown that race often influences the outcomes of interactions with police.

**Question #4**
Within the age groups with the identified age groups that have inconsistencies in searches and contraband found, how does race break down? Are people of a certain race being searched without contraband found at a higher rate?

**Answer**

```{r}
searches_by_age_and_race <- age_grouped_san_diego %>%
  drop_na(subject_age)%>%
  filter(search_conducted == "TRUE")%>%
  group_by(age_group, subject_race) %>%
  summarise(count=n())%>%
  arrange(desc(count)) %>%
  rename(searches = count)

contraband_by_age_and_race <- age_grouped_san_diego %>%
  drop_na(subject_age)%>%
  filter(contraband_found == "TRUE")%>%
  group_by(age_group, subject_race) %>%
  summarise(count=n())%>%
  arrange(desc(count)) %>%
  rename(contraband = count)

race_percentage_search_contraband <- contraband_by_age_and_race %>%
    inner_join(searches_by_age_and_race, by=c("age_group","subject_race"))

race_percentage_search_contraband <- race_percentage_search_contraband %>%
  mutate(percent_contraband=(contraband/searches)*100) %>%
  arrange(desc(percent_contraband))

stops_by_age_and_race <- age_grouped_san_diego %>%
  drop_na(subject_age, subject_race)%>%
  group_by(age_group, subject_race) %>%
  summarise(count=n())%>%
  arrange(desc(count)) %>%
  rename(stops = count)

stops_percentage_searches <- stops_by_age_and_race %>%
  inner_join(searches_by_age_and_race, by=c("age_group","subject_race"))

stops_percentage_searches <- stops_percentage_searches %>%
  mutate(percent_searched=(searches/stops)*100) %>%
  arrange(desc(percent_searched))

race_percentage_stops_search_contraband <- race_percentage_search_contraband %>%
  inner_join(stops_percentage_searches, by=c("age_group","subject_race","searches"))

```

**Question 5**
Where are these stops with the identified inconsistencies and race happening? Is there a concentration in a certain area? Does that area have high populations of a certain race?

**Answer**

```{r}
mega_stops<-age_grouped_san_diego%>%
  drop_na(subject_age,)%>%
    drop_na(subject_race)%>%
  group_by(age_group, subject_race,service_area) %>%
  summarise(count=n())%>%
   arrange(desc(count))

mega_searches <- age_grouped_san_diego %>%
  drop_na(subject_age)%>%
  drop_na(subject_race)%>%
  filter(search_conducted == "TRUE")%>%
  group_by(age_group, subject_race, service_area) %>%
  summarise(count=n())%>%
  arrange(desc(count))

mega_seaches_mega_stops<-mega_stops%>%
  inner_join(mega_searches, by=c("age_group","subject_race", "service_area"))%>%
  rename(stops=count.x, searches=count.y)

percent_mega_seaches_mega_stops <- mega_seaches_mega_stops %>%
  mutate(percent_searched=(searches/stops)*100) %>%
  arrange(desc(percent_searched))

```