---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** 
Most loans are in the $10,000-$50,000 category. The fewest are in the above $1,000,000 category and the $500,000-$1,000,000 category.

```{r}
wv_loan_amount_categories <- west_virginia_ppp %>%
mutate(
  amount_category = case_when(
    amount < 10000 ~ 'under_10000',
    amount >= 10000 & amount < 50000 ~ '10000_to_50000',
    amount >= 50000 & amount < 100000 ~ '50000_to_100000',
    amount >= 100000 & amount < 500000 ~ '100000_to_500000',
    amount >= 500000 & amount < 1000000 ~ '500000_to_1000000',
    amount >= 1000000 ~ '1000000'
  )
) %>%
  group_by(amount_category) %>%
  summarise(
    count_loans = n()
  )

```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Davis Trust Company has 12 loans with undisbursed amounts, and Miners and Merchants bank has 1 loan with an undisbursed amount, for 13 total loans.

```{r}
west_virginia_ppp %>% filter(undisbursed_amount > 0)

```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** 
The loan might not be paid out because the lender could have run out of money, or something in the money transfer might be wrong, like a bank account number, or the business could not be communicating with the lender. I would want to know why one lender has the overwhelming majority of undisbursed loans, and if they have paid out some of the total money. If the lender has run out of money, I would want to know what the business's options are for still accessing funds.


**Q4** Which month in 2021 had the most loans approved?
**A4** March

```{r}
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  filter(Month>"2020-12-01")
```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** 58 loans went to West Virginia applicants. J. Austin Heating and Cooling Inc. had the earliest approval on April 8, 2020.

```{r}
west_virginia_ppp %>%
   filter(jobs_retained == 500) %>%
  arrange(date_approved)

```
