---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan.27

data is case sensitive

the command to set a working directory is setwd()
```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```
you can run individual commands in markdown notebooks
library-ing the tidyverse is very important

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("tidycensus")
```

reading rds is a tidyverse function, you have to load it first

saving as a variable, containing the results of the code
```{r}
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans=n()
  )
```

Mutation code
```{r}
mutate(name_of_column = value)
```

Select code, only shows the columns you ask for, but don't make it permanent
```{r}
  select(loan_number, amount, payroll_proceed)
```

Filtering
```{r}
maryland_ppp %>%
  filter(name_of_column == 'word')

  filter(is.na(forgiveness_date))
```

cleaned_wv_ppp_loans <- wv_ppp_loans %>%
  clean_names() %>%
  distinct() %>%
  mutate(city = str_to_title(city))
  
  making dataframe
```{r}
  write_csv(college_park_loans, "data/college_park_loans.csv")

```


rvest
read_html()- give url, reads the HTML
html_text()- gives text inside of a tag
html_table()- given some html, finds all the <table>
html_element()- given some html, supply a tag like p or ul OR xpath expression
```{r}
my_stuff <- url %>%
  read_html() %>%
  html_element('p b') #this finds all bold tags in p tags
```

```{r}
age_under_18 <- san_diego_stops %>%
  filter(subject_age < 18) %>%
  nrow()

age_18_24 <- san_diego_stops %>%
  filter(subject_age >=18, subject_age<25) %>%
  nrow() 
  
age_25_34 <- san_diego_stops %>%
  filter(subject_age >=25, subject_age<35) %>%
  nrow()

age_35_44 <- san_diego_stops %>%
  filter(subject_age >=35, subject_age<45) %>%
  nrow()

age_45_54 <- san_diego_stops %>%
  filter(subject_age >=45, subject_age<55) %>%
  nrow()

age_55_64 <- san_diego_stops %>%
  filter(subject_age >=55, subject_age<65) %>%
  nrow()

age_65_and_up <- san_diego_stops %>%
  filter(subject_age >=65) %>%
  nrow()
```