---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan.27

data is case sensitive

the command to set a working directory is setwd()
```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```
you can run individual commands in markdown notebooks
library-ing the tidyverse is very important

```{r}
install.packages("tidyverse")
library(tidyverse)
```

reading rds is a tidyverse function, you have to load it first

saving as a variable, containing the results of the code
```{r}
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans=n()
  )
```

Mutation code
```{r}
mutate(name_of_column = value)
```

Select code, only shows the columns you ask for, but don't make it permanent
```{r}
  select(loan_number, amount, payroll_proceed)
```

Filtering
```{r}
maryland_ppp %>%
  filter(name_of_column == 'word')

  filter(is.na(forgiveness_date))
```

cleaned_wv_ppp_loans <- wv_ppp_loans %>%
  clean_names() %>%
  distinct() %>%
  mutate(city = str_to_title(city))
  
  
  