---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidycensus)
library(tidyverse)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("7fa645db37d4746697c8b06053a048d8b9708aa9", install=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
acs_2019 <- load_variables(2019, "acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
ppp_applications <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1**
Baltimore City

```{r}
v19 <- load_variables(2019, "acs5", cache = TRUE)

md_2019_age <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2019)

md_2019_age %>% 
  arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2**
Talbot County and Worchester County

```{r}
md_2019_over_50 <- md_2019_age %>%
  filter(estimate > "50")

```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3**
Prince George's County, Montgomery County, Caroline County, Frederick County, and Talbot County have the highest percentages of 5 to 17 year olds who speak Spanish at home

```{r}
md_5_17 <- get_acs(geography = "county",
              variables = c(agerange = "B16007_002"),
              state = "MD",
              year = 2019)

md_5_17_spanish <- get_acs(geography = "county",
              variables = c(spanishspeaking = "B16007_004"),
              state = "MD",
              year = 2019)

md_5_17 %>%
  inner_join(md_5_17_spanish, by=c("GEOID", "NAME")) %>%
  rename(estimate_5to17=estimate.x, estimate_spanish=estimate.y) %>%
  select(-variable.x,-variable.y,-moe.x,-moe.y) %>%
  mutate(percent_spanish =(estimate_spanish/estimate_5to17)*100) %>%
  arrange(desc(percent_spanish))
```
	

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**
PG County and Montgomery County both have very large populations and are in the DC metro area, so it is not surprising that they have a large percentage. Caroline and Talbot are more interesting because they both have small young populations and Talbot has a median age over 50, so they have unusually large percentages. Garrett County has a very small young population, and a very small Spanish speaking young populations.

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5**
60620 is the zip code with the most removed PPP applications, and it has a 95.7% Black population

```{r}
pop_by_race <- get_acs(geography = "zcta",
              variables = c(pop_race = "B02001_001"),
              year = 2019)

pop_black <- get_acs(geography = "zcta",
              variables = c(pop_black = "B02001_003"),
              year = 2019)

pop_race_with_black <- pop_by_race %>%
  inner_join(pop_black, by=c("GEOID", "NAME")) %>%
  rename(estimate_race=estimate.x, estimate_black=estimate.y) %>%
  select(-variable.x,-variable.y,-moe.x,-moe.y) %>%
  mutate(percent_black =(estimate_black/estimate_race)*100) %>%
  arrange(desc(percent_black))

ppp_applications <- ppp_applications %>%
  mutate(zip5 = str_sub(zip, 1, 5)) 

top_zips <- ppp_applications %>%
  group_by(zip5) %>%
  summarise(total=n()) %>%
  arrange(desc(total))

pop_race_with_zips <- top_zips %>% left_join(pop_race_with_black, by=c('zip5'='GEOID'))
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**
Of the top 10 zip codes with the most removed, six have a percent of Black population above 90% and a seventh is 89%. All but one of the top 10 are located in or around Chicago in Cook County (the South side of the Chicago area), with the exception being a 91% Black zip code in Atlanta. Clearly, an alarmingly large amount of PPP loans were removed in Chicago, specifically the overwhelmingly Black areas. Other notable zip codes are in majority Black areas around Atlanta. 

**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7**
It would be useful to see the race column of the applicants, if there is sufficient data in that category, which there may not be. I would also find it useful to see the lenders, I would assume that these loans were through the same or related lenders if they are all in the same area. NAICS codes would be interesting to look into, as well as the business type category, to see if specific industries or business types may have been targeted.
