---
title: "Reverse Engineering Project"
author: "Destiny Herbers"
date: "March 4, 2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Nearly 250 women have been fatally shot by police since 2015 (Washington Post) https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/ 

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Of men killed, 57 percent were armed with a gun and of women, 44 percent. In other cases, women were armed with knives, cars, toy weapons, hammers and hatchets.
* **Analysis summary**: We brought the data into open refine to make a new column and get all of the weapons that include a gun to sort. We have the code to get the total killed men and total killed men who were armed with a gun, but we are stuck with how to calculate this percentage within R because there are no number columns to use in a formula, and we're not sure how to mutate for what we need.

```{r}
police_shootings_men <- fatal_police_shootings %>%
  filter(gender == "M") %>%
  group_by(armed_clean) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
  


# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. 
* **Analysis summary**: We ended up with similar numbers, but more total women, with 15 white women, 7 Black women, 5 Hispanic women, and 5 whose race were unknown. This might be because the Post considered women as only adults, but we do not know exactly what values they used to filter.

```{r}
fatal_police_shootings %>%
  filter(armed == "unarmed", gender == "F") %>%
  group_by(armed, gender, race) %>%
  summarise(count=n())


# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-