---
title: "Reverse Engineering Project"
author: "Destiny Herbers"
date: "March 4, 2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Nearly 250 women have been fatally shot by police since 2015 (Washington Post) https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/ 

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Of men killed, 57 percent were armed with a gun and of women, 44 percent. In other cases, women were armed with knives, cars, toy weapons, hammers and hatchets.
* **Analysis summary**: We brought the data into Open Refine to make a new column and get all of the weapons that included a gun to sort. When we calculated the percentages, we got roughly 58.9% of men and 45.3% of women, which is very close to the numbers in the story. Our count may be higher because the data may have been updated to include more instances after the Post made their calculations and published the story. 

```{r}
all_police_shootings_by_gender <- fatal_police_shootings %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

police_shootings_men <- all_police_shootings_by_gender %>% 
  filter(gender == 'M')

police_shootings_women <- all_police_shootings_by_gender %>% 
  filter(gender == 'F')

gun_police_shootings_by_gender <- fatal_police_shootings %>%
  filter(armed_clean == "gun") %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

all_shootings_men_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'M')

all_shootings_women_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'F')

men_with_gun_percentage <- ((4012/6814)*100)
women_with_gun_percentage <- ((145/320)*100)

# Display results of code below this codeblock
men_with_gun_percentage
women_with_gun_percentage
```

### Sentence 2

* **Sentence text**: Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. 
* **Analysis summary**: We ended up with similar numbers, but more total women, with 15 white women, 7 Black women, 5 Hispanic women, and 5 whose race were unknown. This might be because the Post considered women as only adults, but we do not know exactly what values they used to filter. It also may be because the Post analyzed this data before we did, so more cases could have been added since then.

```{r}
shootings_women_race <- fatal_police_shootings %>%
  filter(armed == "unarmed", gender == "F") %>%
  group_by(armed, gender, race) %>%
  summarise(count=n())


# Display results of code below this codeblock
shootings_women_race
```

### Sentence 3

* **Sentence text**: Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known.
* **Analysis summary**: We removed all of the rows with N/A in the race column to only have situations where race was known and got a total number. Then we filtered for Black women and got that total number. We calculated a percentage with both numbers and got 0.9%, which is less than 1 percent, and replicates the results found by the post. Once again it is possible that our total was higher if the dataset has been updated since the story was published, but it is still technically correct.

```{r}
# Put code to reverse engineer sentence here
shootings_race_known <- fatal_police_shootings %>%
  drop_na(race)

shootings_Black_women <- shootings_race_known %>%
  filter(gender == "F", race == "B") %>%
  summarise(count=n())

shootings_race_known %>%
    summarise(count=n())

percent_Black_women <- ((54/5941)*100)
  
# Display results of code below this codeblock
shootings_race_known
shootings_Black_women
percent_Black_women
```


***A note to you, Mr Willis*** 
In the sentences we had originally chosen, the Post used information that isn't in the data we were given, specifically about the mental health status of the people who were killed and if they were killed in their homes. This data isn't in the rest of the GitHub repository either, so we went back to the story to find sentences that we could solve with the categories in the data set. We wanted more variance, but we were limited. 
***End note***

### Sentence 4

* **Sentence text**: "Black women... account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths."
* **Analysis summary**: For the percentage of women killed that were Black, we chose to calculate with only instances where race was known because they didn't specify in this sentence. We got 19.3%, which is within the margin of all of our calculations, so we believe we have replicated it. For the percentage of the unarmed deaths, we assumed that they were still referring to unarmed women, though it was not clear. We got 25.9%, which is outside of our normal error. These numbers could be different because they data was changed after publication, or because they used totals where race was not known.

```{r}
all_women_shootings <- shootings_race_known %>%
  filter(gender == "F") %>%
  summarise(count=n())

shootings_Black_women

percent_Black_of_women <- ((shootings_Black_women/all_women_shootings)*100)

unarmed_deaths_women <- shootings_race_known %>%
  filter(gender == "F", armed == "unarmed") %>%
  summarise(count=n())

Black_women_unarmed <- shootings_race_known %>%
  filter(gender == "F", race == "B", armed == "unarmed") %>%
  summarise(count=n())

percent_Black_unarmed_women <- ((Black_women_unarmed/unarmed_deaths_women)*100)

# Display results of code below this codeblock
percent_Black_of_women
percent_Black_unarmed_women
```

### Sentence 5

* **Sentence text**: "Black men... make up 27 percent of the men shot and 36 percent of the unarmed deaths."
* **Analysis summary**: For the percentage of men shot that were Black, we got 27%, which replicates their findings. Again, we assumed that they were working with shootings where race is known. The percentage of unarmed men who were Black in our calculation is 34%, which again is about 2% less than the Post's total. This consistency is interesting, because it seems unlikely that if more deaths were added post-publication, an equal number of unarmed non-Black men and women were killed. It is possible that because of the vague wording used in both sentences, the Post could have calculated their total differently than we did.

```{r}
all_men_shootings <- shootings_race_known %>%
  filter(gender == "M") %>%
  summarise(count=n())

shootings_Black_men <- shootings_race_known %>%
  filter(gender == "M", race == "B") %>%
  summarise(count=n())

percent_Black_of_men <- ((shootings_Black_men/all_men_shootings)*100)

unarmed_deaths_men <- shootings_race_known %>%
  filter(gender == "M", armed == "unarmed") %>%
  summarise(count=n())

Black_men_unarmed <- shootings_race_known %>%
  filter(gender == "M", race == "B", armed == "unarmed") %>%
  summarise(count=n())

percent_Black_unarmed_men <- ((Black_men_unarmed/unarmed_deaths_men)*100)

# Display results of code below this codeblock
percent_Black_of_men
percent_Black_unarmed_men
```

-30-