---
title: "Reverse Engineering Project"
author: "Destiny Herbers"
date: "March 4, 2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Nearly 250 women have been fatally shot by police since 2015 (Washington Post) https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/ 

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Of men killed, 57 percent were armed with a gun and of women, 44 percent. In other cases, women were armed with knives, cars, toy weapons, hammers and hatchets.
* **Analysis summary**: We brought the data into open refine to make a new column and get all of the weapons that included a gun to sort. When we calculated the percentages, we got roughly 58.9% of men and 45.3% of women, which is very close to the numbers in the story. Our count may be higher because the data may have been updated to include more instances after the Post made their calculations and published the story. 

```{r}
all_police_shootings_by_gender <- fatal_police_shootings %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

police_shootings_men <- all_police_shootings_by_gender %>% 
  filter(gender == 'M')

police_shootings_women <- all_police_shootings_by_gender %>% 
  filter(gender == 'F')

gun_police_shootings_by_gender <- fatal_police_shootings %>%
  filter(armed_clean == "gun") %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

all_shootings_men_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'M')

all_shootings_women_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'F')

men_with_gun_percentage <- ((4012/6814)*100)
women_with_gun_percentage <- ((145/320)*100)

# Display results of code below this codeblock
men_with_gun_percentage
women_with_gun_percentage
```

### Sentence 2

* **Sentence text**: Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. 
* **Analysis summary**: We ended up with similar numbers, but more total women, with 15 white women, 7 Black women, 5 Hispanic women, and 5 whose race were unknown. This might be because the Post considered women as only adults, but we do not know exactly what values they used to filter. It also may be because the Post analyzed this data before we did, so more cases could have been added since then.

```{r}
shootings_women_race <- fatal_police_shootings %>%
  filter(armed == "unarmed", gender == "F") %>%
  group_by(armed, gender, race) %>%
  summarise(count=n())


# Display results of code below this codeblock
shootings_women_race
```

### Sentence 3

* **Sentence text**: Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known.
* **Analysis summary**: We removed all of the rows with N/A in the race column to only have situations where race was known and got a total number. Then we filtered for Black women and got that total number. We calculated a percentage with both numbers and got 0.9%, which is less than 1 percent, and replicates the results found by the post. Once again it is possible that our total was higher if the dataset has been updated since the story was published, but it is still technically correct.

```{r}
# Put code to reverse engineer sentence here
shootings_race_known <- fatal_police_shootings %>%
  drop_na(race)

shootings_Black_women <- shootings_race_known %>%
  filter(gender == "F", race == "B") %>%
  summarise(count=n())

shootings_race_known %>%
    summarise(count=n())

percent_Black_women <- ((54/5941)*100)
  
# Display results of code below this codeblock
shootings_race_known
shootings_Black_women
percent_Black_women
```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-