---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data.

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** 
Rows: 33,336 
Columns: 58

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

Use the pre-lab and your reference notebook as examples.

**Q2** Which West Virginia county got the most loans? How many?
**A2** 
Kanawha, 4273

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))

```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** 
Palm Beach

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
    arrange(count_loans)

```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** Kanawha County is the most populous county in West Virginia and has the state capital, so it makes sense that they would have the most businesses applying for loans. Palm Beach isn't in West Virginia, and neither is Allegheny, so I'm guessing that this is a trick question and the data is wrong, maybe a couple businesses marked down the wrong state.


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** Greenbrier Hotel Corporation got the largest PPP loan for $8,912,202. The corporation owns and operates a luxury resort in White Sulphur Springs. The loan retained 500 jobs.

```{r}
wv_ppp_loans %>%
  arrange(desc(amount))

```

**Q6** What West Virginia politician (with his family) owns the company identified in question 5? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** Jim Justice, governor of West Virginia, owns the hotel and his family hold most of the top positions in the company. My source is West Virginia Metro News: https://wvmetronews.com/2021/07/25/in-swirl-of-financial-turmoil-greenbrier-hotel-corp-is-hit-with-million-dollar-irs-lien/ and the Washington Post: https://www.washingtonpost.com/lifestyle/magazine/wva-billionaire-jim-justices-mission-to-restore-the-greenbrier-resort/2011/02/09/AB5NO5M_story.html

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** The smallest loan received is only $165, so it might be a good story to find out what companies are applying for such little money and how they use it. The average loan amount is also only $78,016 and the median amount is even lower at $20832, where as the maximum is almost 9 million, and went to a business owned by the governor. I think it could be interesting to find out how the businesses that received the highest amount of money compare to the others.
```{r}
wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) 
saveRDS(object,file="wv_summary.rds")
```
```{r}
wv_summary <- read_rds("wv_summary.rds")
```