---
title: "state_profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tigris)
library(tidycensus)
library(ggrepel)
library(sf)
```

```{r}
maine_ppp <- read_csv("data/maine_ppp.csv") %>% clean_names()
```

## Initial Exploration

**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

```{r}
maine_ppp_clean <- maine_ppp %>%
  distinct() %>%
  unique() %>%
  mutate(zip = str_sub(zip, 1, 5))
```

```{r}
maine_by_county <- maine_ppp_clean %>%
  drop_na(project_county_name) %>%
  group_by(project_county_name) %>%
  summarise(count=n())%>%
  arrange(desc(count))

maine_by_zip <- maine_ppp_clean %>%
  drop_na(zip) %>%
  group_by(zip) %>%
  summarise(count=n())%>%
  arrange(desc(count))

maine_by_lender <- maine_ppp_clean %>%
  group_by(lender) %>%
  summarise(count=n())%>%
  arrange(desc(count))

maine_bsb_loans <- maine_ppp_clean %>%
  filter(lender == "Bangor Savings Bank") %>%
  distinct() %>%
  group_by(project_county_name) %>%
  summarise(count=n())%>%
  arrange(desc(count))

cumberland_loans <- maine_ppp_clean %>%
  filter(project_county_name == "CUMBERLAND") %>%
  nrow()

percent_bsb_cumberland <- (2387/12950)*100

maine_by_loan_status <- maine_ppp_clean %>%
  group_by(loan_status) %>%
  summarise(count=n())%>%
  arrange(desc(count))

exemption_4 <- maine_ppp_clean %>%
  filter(loan_status == "Exemption 4") %>%
  group_by(undisbursed_amount) %>%
  summarise(count=n())%>%
  arrange(desc(count))
```

**A1.** 
Maine had 47,407 PPP loan applications, I cleaned the data for distinct and unique entries, though there were some repeated names in Open Refine, there are no repeated ID numbers, so I believe that those are unique loans. I also cleaned up the city column and the county column and standardized the zipcode column to 5 digit entries.

Cumberland County has the most loans in Maine by far, 12950, followed by York County with 6483, Penobscot with 4083, Hancock with 3953, and Kennebec with 2872. Cumberland County is the most populous, and has Portland, the largest city in Maine, so the number of loans makes sense. York County is directly below Cumberland, with the outskirts of Portland and many coastal tourist towns. Kennebec is surprisingly low considering it features the capital city Augusta, though Augusta is the third least populous state capital. The individual zip code with the highest total amount of loans was 04101 with 2073 loans, which encompasses Portland, followed by 04401 with 1633 in Penobscot which has the city Bangor. Bangor Savings Bank was actually the lender that serviced the most loans in Maine, 6939 total, and is the largest bank in Maine with 54 locations. Most of Bangor Savings Bank's loans were in Cumberland County and Penobscot County, with 2387 and 1169 respectively. Bangor Savings Bank serviced roughly 18% of the PPP loans out of Cumberland County. 43953 of the loans are paid in full, and 3454 are listed as Exemption 4, which I researched and found to most likely be referencing FOIA Exemption 4, which relates to trade secrets. I filtered to look specifically at Exemption 4 loans, and they had all other information present, and all but three loans had no undisbursed amount, so I am unsure how the payment status is FOIA exempt information.

## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure. Describe the most interesting or newsworthy findings based on your exploration.

```{r}
#state population of Maine is 1335492
#state per capita is (47407/1335492)*100000 = 3550
maine_pop <- get_acs (geography = "state",
                           variables=c("B01001_001"),
                           year=2019,
                           state = 'ME')

#by zip with state difference
maine_pop_zip <- get_acs (geography = "zcta",
                           variables=c("B01001_001"),
                           year=2019,
                           state = 'ME',
                           geometry = TRUE) %>%
                           rename(zip=NAME)

maine_pop_zip <- maine_pop_zip %>%
        mutate(zip = str_remove_all(zip, "ZCTA5 ")) %>%
        select(-GEOID,-variable,-moe)

maine_ppp_zip_pop <- maine_pop_zip %>%
  inner_join(maine_by_zip, by=c("zip")) %>%
  mutate(loans_per_100k = (count/estimate)*100000) %>%
  mutate(state_difference = (3550 - loans_per_100k))

ggplot() +
  geom_sf(data=maine_ppp_zip_pop, aes(fill=state_difference)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

#by county
maine_pop_county <- get_acs (geography = "county",
                           variables=c("B01001_001"),
                           year=2019,
                           state = 'ME',
                           geometry = TRUE) %>%
                    mutate(NAME = trimws(toupper(NAME))) %>%
                    mutate(NAME = str_remove_all(NAME, " COUNTY, MAINE"))

maine_pop_county <- maine_pop_county %>%
  rename(project_county_name = NAME) 

maine_ppp_county_pop <- maine_pop_county %>%
  inner_join(maine_by_county, by=c("project_county_name")) %>%
  select(-GEOID,-variable,-moe) %>%
  mutate(loans_per_100k = (count/estimate)*100000)

ggplot() +
  geom_sf(data=maine_ppp_county_pop, aes(fill=loans_per_100k)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

```

**A2.** 
The findings seem pretty consistent with my prior findings, the zip code that contains Portland has a higher per capita loan amount than the state overall by far, and a lot of areas in the state had negative values. It's interesting that there are some zip codes in the North that have a pretty high per capita difference from the state, they could be worth further investigation. The county breakdowns are more interesting, I expected Cumberland and York counties to be in the top tier, but it looks like Hancock and Knox counties have the highest rates of loans per 100k people. This is surprising because Knox County had not been notable so far.

## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below.

```{r}
maine_by_lender <- maine_ppp_clean %>%
  group_by(lender) %>%
  summarise(count=n(), total_amount=sum(amount))%>%
  mutate(average_amount = total_amount/count) %>%
  arrange(desc(count))

#all loans have 0 undisbursed, other than 3 N/A
maine_undisbursed <- maine_ppp_clean %>%
  group_by(undisbursed_amount) %>%
  summarise(count=n())

#lender and amount relationship
maine_amount_lender <- maine_ppp_clean %>%
  group_by(lender, amount) %>%
  summarise(count=n())%>%
  arrange(desc(count))

ggplot() +
  geom_point(data=maine_amount_lender, aes(x=amount, y=count)) 

#amount breakdown
lender_amount_groups <-maine_amount_lender %>%
mutate(
  amount_groups = case_when(
    amount < 1000 ~ 'under_1k',
    amount >= 1000 & amount < 10000 ~ '1k_10k',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 500000 ~ '100k_500k',
    amount >= 500000 & amount < 1000000 ~ '500k_1m',
    amount >= 1000000 & amount < 5000000 ~ '1m_5m',
    amount >= 5000000 & amount < 8000000 ~ '5m_8m',
    amount >= 8000000 ~'8m+'
))

ggplot() +
  geom_point(data=lender_amount_groups, aes(x=amount_groups, y=count))

```

**A3.** 
Lenders tend to give the most loans in the 10k to 50k category, and looking at the data frame shows that those loans are mostly for $20800 or $20833, which makes me wonder if there's a reason that a bunch of banks gave loans for that specific amount. There were a few loans above $1 million, and most of the banks that lended for those loans only did one in that value range, so it seems like there wasn't one bank that was doing large values. There are also many loans under $1,000, which is interesting because that is probably not a significant amount of money for most businesses.

## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

```{r}
census_naics <- read_csv("/Users/destinyherbers/Downloads/cbp19co.txt") %>% clean_names()

all_fips <- fips_codes %>% as_tibble()
maine_fips <- all_fips %>%
  filter(state == "ME")

maine_fips <- maine_fips %>%
  mutate(fips=str_c(state_code, county_code)) %>%
  mutate(county = trimws(toupper(county))) %>%
  mutate(county = str_remove_all(county, " COUNTY"))

maine_ppp_fips <- maine_fips %>%
  left_join(maine_ppp_clean, by=c("county" = "project_county_name"))

maine_naics_totals <- maine_ppp_fips %>%
  group_by(naics_code) %>%
  summarise(count=n())%>%
  arrange(desc(count)) %>%
  head(10)

clean_census <- census_naics %>%
  filter(!str_detect(naics,"-")) %>%
  filter(!str_detect(naics, "/")) %>%
  filter(str_length(naics) == 6) %>%
  filter(fipstate == 23) %>%
  mutate(fips = str_c(fipstate, fipscty))

maine_census_naics_totals <- clean_census %>%
  group_by(naics) %>%
  summarise(businesses = sum(est))%>%
  arrange(desc(businesses)) %>%
  head(10)

no_self_or_sole <- maine_ppp_fips %>%
  filter(business_type != "Sole Proprietorship", business_type != "Self-Employed Individuals") %>%
    group_by(naics_code) %>%
    summarise(count=n())%>%
    arrange(desc(count)) %>%
    head(10)

#812112, beauty salon analysis
maine_salon_ppp <- maine_ppp_fips %>% 
  filter(naics_code == "812112") %>% 
  group_by(county, fips) %>% 
  summarise(total=n()) %>% 
  arrange(desc(total))

census_salon_maine <- clean_census %>%
  filter(naics == "812112") %>% 
  group_by(fips) %>% 
  summarise(businesses = sum(est))%>%
  mutate(businesses = replace_na(businesses, 0)) %>%
  arrange(desc(businesses)) 

maine_salons_combined <- maine_salon_ppp %>%
  inner_join(census_salon_maine, by=c("fips")) %>%
  mutate(difference = total-businesses)

write.csv(maine_salons_combined, file = "maine_salons_combined.csv")
```

**A4.** 
Maine overall has the most businesses in naics code 722511, which is full service restaurants with 1483, followed by 722513, which is limited service restaurants with 998. Those categories also show up in the ppp loan data, full service restaurants have the second highest number of loans with 2057, and limited service restaurants are the eighth highest with 732 loans. The industry with the most ppp loans was 114112, businesses handling shellfish, with 2815 loan applications. This industry does not appear in the top census data. With self employed individuals and sole proprietorships removed, the shellfish industry falls to number five, and the top spots match pretty well with the census data. 

Visualization: https://datawrapper.dwcdn.net/r3cJw/1/

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.**
Most of Maine's data is pretty straightforward, there were very few high-value loans and for the most part the loan density was pretty equal to population density, centered around the significant cities. It is notable that Knox County has a very high rate of loan applications per 100 thousand people because it's a very small county with few towns and mostly a ton of coastland, so I would not expect to see so many businesses applying for loans. The high amount of loans could be from self-employed fishermen or an industry like that. To report about Knox County, I would need to explore the loans from that county, what addresses they come from and what naics codes they reported to see if there's a logical explanation for the density. There also were surprisingly few loans in the center of the state, particularly around the capital, Augusta, which could be interesting to report about. As a minimum story, I think that it's interesting that the loans are concentrated around the coast and the northern border of Maine, and it would be cool to contrast that with business and population densities. The maximum story would probably involve investigating the small coastal counties that have unexpectedly large loan application per capita rates, and to report out what drove all of those applications. If it is the shellfish industry, which I suspect it might be from the previous question, it would be cool to dive deeper into how the seafood workers were impacted by the pandemic.
